Recurrent Deep Learning Models and its applications

### 2025-03-29 paper2/main.tex
fixed bug

\subsubsection{RNN encoder decoder model}
RNN Encoder-Decoderis a RNN based architecture which commonly used for handle sequence to sequence problem, for example machine translation, summarization, and question answering. The designed of this architecture is good for handle the input sequence and the output sequence is not the same length, leveraging the sequential nature of data. 
\\[1ex]
In this architecture, there are two main components, the first component is rnn encoder, the usage of the encoder is to encode the input sequence into a fixed length vector. And the decoder is to decode the vector into the target sequence. 
\\[1ex]
\textbf{Encoder}
\\[1ex]
The encoder is a RNN based block, which processes the input sequence step by step and calculate the hidden state representation for each input token. Let the input sequence be $X = [x_1, x_2, \cdots, x_T]$ where $T$ is the length of the input sequence, and $x_t$ is the input at time step $t$. We use the below formula to computes the hidden states $h_t$.
\\[1ex]
For the Vanilla RNN encoder
\begin{equation}
    h_t = f_{enc}(x_t, h_{t-1})
\end{equation}

@@

>>>>

\subsubsection{RNN encoder decoder model}
\begin{figure}[!htb]
    \centering
    \includegraphics[width=1\textwidth]{../Pic/RNN_encoder_decoder.jpg} % Replace "example-image" with your image file
\end{figure}
RNN Encoder-Decoderis a RNN based architecture which commonly used for handle sequence to sequence problem, for example machine translation, summarization, and question answering. The designed of this architecture is good for handle the input sequence and the output sequence is not the same length, leveraging the sequential nature of data. 
\\[1ex]
In this architecture, there are two main components, the first component is rnn encoder, the usage of the encoder is to encode the input sequence into a fixed length vector. And the decoder is to decode the vector into the target sequence. 
\\[1ex]
\textbf{Encoder}
\\[1ex]
The encoder is a RNN based block, which processes the input sequence step by step and calculate the hidden state representation for each input token. Let the input sequence be $X = [x_1, x_2, \cdots, x_T]$ where $T$ is the length of the input sequence, and $x_t$ is the input at time step $t$. We use the below formula to computes the hidden states $h_t$.
\\[1ex]
For the Vanilla RNN encoder
\begin{equation}
    h_t = f_{enc}(x_t, h_{t-1})
\end{equation}

@@

\subsubsection{Attention Mechanism}
% Vaswani et al. introduces the Transformer architecture, which replaces RNN model with self attention. In this section, I want to review the attention machanism. \parencite{vaswani2023attentionneed}. Attention machanism is good for handle the sequence-to-sequence tasks such as machine translation, text summarization and image captioning. Bahdanau et al. \parencite{bahdanau2016neuralmachinetranslationjointly} introduces the attention mechanism in the machine translation task. At each decoding step $t'$, the model model will compute the attention scores $\alpha_{t',t}$ that represent how important of the current hidden state $h_t$ to the current decoder state $s_{t'}$. We have different attention formula.

\textbf{Bahdanau (Additive) Attention}
\\[1ex]
% This method proposed by \parencite{bahdanau2016neuralmachinetranslationjointly}, in which this attention uses a feedforward network to compute the attention scores. 
\\[1ex]
\begin{equation}
    e_{t',t} = v_a^Ttanh(W_as_{t'-1}+U_ah_t)
\end{equation}                    
\textbf{Luong (Multiplicative) Attention}
\\[1ex]
% This method is introduced by Luong et al \parencite{luong2015effectiveapproachesattentionbasedneural}. This method uses the dot product to compute the attention scores.
\begin{equation}
    e_{t',t} = s_{t'-1}^Th_t
\end{equation}
\textbf{Scaled Dot-Product Attention}
\\[1ex]
